Article Title

### 2013-08-13 sections/Introduction.tex
minor

Neural correlations do not come with ready or unambiguous mechanistic interpretations.  Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms.  These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

@@

With this study, we pursue two goals: 
\begin{enumerate}[\qquad A.\;\;]
	\item Devise better estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
\end{enumerate}

>>>>

However, neural correlations do not come with ready or unambiguous mechanistic interpretations.  Theoretical and simulation studies have shown that neural correlations may arise at various temporal scales from combinations of multiple underlying mechanisms.  These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

@@

With this study, we pursue two related goals: 
\begin{enumerate}[\qquad A.\;\;]
	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
\end{enumerate}

### 2013-08-13 sections/Theory.tex
minor

\subsection{Overview}
Given a finite sample, the \emph{sample covariance matrix} $\hat\Sigma_0$ is the usual estimator of the \emph{true covariance matrix} $\Sigma$ of the underlying random process.  $\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.  More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a different, low-dimensional parametric \emph{target estimate} $\hat T$. 
Target estimates are biased approximations because they cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.

A regularized estimate requires the following three choices: 
\begin{enumerate}[  1. ]
\item define a set of low-dimensional target estimates
\item given the data, select the optimal target estimate from the set 
\item given the data, select the optimal amount of biasing toward the target estimate
\end{enumerate}
The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.  Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.  The perplexing fact that estimates can be improved by biasing toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.

@@

The usual unbiased estimate, the sample covariance matrix, rescales the plug-in estimator to compensate for the bias
\begin{equation}
\hat\Sigma_0 = \frac n {n - d} \hat\Sigma \equiv \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
\end{equation}
When observations are independent, then $d=1$.  But when nearby samples are correlated, the estimate of the mean absorbs more of the variance, becomes \emph{overdispersed}. As a result the covariances become underestimated. Then we can use $d=cn$ where $c>1$ is called the \emph{overdispersion coefficient}, which may be estimated from the data.

@@

When the loss function is the error sum of squares, which, for matrices, is called the Frobenius norm, 
\begin{equation}
\mathcal L_e(\hat S,S) = \|\hat S-S\|_F^2 = \Tr\left((\hat S-S)(\hat S-S)^\T\right)
\end{equation}
the variance-bias decomposition is the simple sum
\begin{equation}
r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
\end{equation}

>>>>

\subsection{Overview}
Given the sample $x_1,\ldots,x_n$ of $p\times 1$ observations, the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
\begin{equation}
\hat\Sigma_0 =  \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
\quad\mbox{where}\;
\bar x = \frac 1 n \sum\limits_{i=1}^n 
\end{equation}
where the \emph{index of disperson} $d=1$ when observations are independent or $d>1$ when nearby samples are correlated; the value of $d$ can be estimated from the data.

$\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.  More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a less variable \emph{target estimate} $\hat T$. Target estimates are low-dimensional approximations cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.

\subsection{Target estimates}
A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.

A regularized estimate requires the following three choices: 
\begin{enumerate}[  1. ]
\item define a set of low-dimensional target estimates
\item given the data, select the optimal target estimate from the set 
\item given the data, select the optimal amount of biasing toward the target estimate
\end{enumerate}
The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.  Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.  The perplexing fact that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.

@@

@@

When the loss function is the mean squared error, which, for matrices, is called the normalized, squared Frobenius norm: 
\begin{equation}
\mathcal L_e(\hat S,S) =\tfrac 1 n \|\hat S-S\|_F^2 = \tfrac 1 n \Tr\left((\hat S-S)(\hat S-S)^\T\right)
\end{equation}
the variance-bias decomposition is the simple sum
\begin{equation}
r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
\end{equation}
